Outbreak Threshold Detection

### 2023-03-03 notebooks/Julia/simulation-detection.qmd
Create separate jumps for each compartment and calculate N, rather than being separate jump Wanted each compartment to be able to be born and die separately, and calculating N ensures that the total number is always correct (rather than potentially drifting from the sum of the compartments), which is important when values depend on N.

```{julia}
X = 450
Y = 50
Z = 0
N = X + Y + Z
u₀ = [X, Y, Z, N]
```

```{julia}
R₀ = 12
γ = 1/8
μ = 1 / (365 * 62.5)
β = R₀ * (γ + μ)
p = (; β, γ, μ)
```

```{julia}
tmax = 100 * 365
δt = 0.001
tspan = (0.0, tmax)
tlength = length(tspan[1]:δt:tspan[2])
Random.seed!(1234);
```

```{julia}
rate1(u, p, t) = p.β * u[1] * u[2] / u[4]  # β*X*Y / N

function affect1!(integrator)
    integrator.u[1] -= 1         # X -> X - 1
    integrator.u[2] += 1         # Y -> Y + 1
    nothing
end

jump = ConstantRateJump(rate1, affect1!)
```

```{julia}
rate2(u, p, t) = p.γ * u[2]         # γ*Y

function affect2!(integrator)
    integrator.u[2] -= 1        # Y -> Y - 1
    integrator.u[3] += 1        # Z -> Z + 1
    nothing
end

jump2 = ConstantRateJump(rate2, affect2!)
```

```{julia}
birth_rate(u, p, t) = p.μ * u[4]    # μ * N

function birth_affect!(integrator)
    integrator.u[1] += 1
    integrator.u[4] += 1
    nothing
end

@@

```{julia}
death_rate(u, p, t) = p.μ * u[4]    # μ * N

function death_affect!(integrator)
    map!(
        i -> ifelse(integrator[i] > 0, integrator.u[i] - 1, integrator.u[i]),
        integrator.u,
        eachindex(integrator.u)
    )

death_jump = ConstantRateJump(death_rate, death_affect!)
```

```{julia}
prob = DiscreteProblem(u₀, tspan, p)
jump_prob = JumpProblem(prob, Direct(), jump, jump2, birth_jump, death_jump)
sol = solve(jump_prob, SSAStepper())

sol_array = Array(sol)
```
```{julia}
f = Figure()
ax = Axis(
    f[1, 1],
    xlabel = "Time",
    ylabel = "Number"
)

>>>>

```{julia}
X = 800
Y = 20
Z = 800
u₀ = [X, Y, Z]
```

```{julia}
R₀ = 16
γ = 1/8
μ = 1 / (365 * 62.5)
β = R₀ * (γ + μ)
p = (; β, γ, μ)
```

```{julia}
tmax = 250
δt = 0.001
tspan = (0.0, tmax)
tlength = length(tspan[1]:δt:tspan[2])
Random.seed!(1234);
```

```{julia}
infection_rate(u, p, t) = p.β * u[1] * u[2] / (u[1] + u[2] + u[3])      # β*X*Y / N

function infection_affect!(integrator)
    integrator.u[1] -= 1         # X -> X - 1
    integrator.u[2] += 1         # Y -> Y + 1
    nothing
end

infection_jump = ConstantRateJump(infection_rate, infection_affect!)
```

```{julia}
recovery_rate(u, p, t) = p.γ * u[2]     # γ*Y

function recovery_affect!(integrator)
    integrator.u[2] -= 1        # Y -> Y - 1
    integrator.u[3] += 1        # Z -> Z + 1
    nothing
end

recovery_jump = ConstantRateJump(recovery_rate, recovery_affect!)
```

```{julia}
birth_rate(u, p, t) = p.μ * (u[1] + u[2] + u[3])        # μ * N

function birth_affect!(integrator)
    integrator.u[1] += 1
    # integrator.u[4] += 1
    nothing
end

@@

```{julia}
X_death_rate(u, p, t) = p.μ * u[1]       # μ * X

function X_death_affect!(integrator)
    integrator.u[1] -= 1
    nothing
end

X_death_jump = ConstantRateJump(X_death_rate, X_death_affect!)
```

```{julia}
Y_death_rate(u, p, t) = p.μ * u[2]       # μ * Y

function Y_death_affect!(integrator)
    integrator.u[2] -= 1
    nothing
end

Y_death_jump = ConstantRateJump(Y_death_rate, Y_death_affect!)
```

```{julia}
Z_death_rate(u, p, t) = p.μ * u[3]       # μ * R

function Z_death_affect!(integrator)
    integrator.u[3] -= 1
    nothing
end

Z_death_jump = ConstantRateJump(Z_death_rate, Z_death_affect!)
```

```{julia}
jumps = [infection_jump, recovery_jump, birth_jump, X_death_jump, Y_death_jump, Z_death_jump]
```

```{julia}
prob = DiscreteProblem(u₀, tspan, p)
jump_prob = JumpProblem(prob, Diredt(), jumps...)
sol = solve(jump_prob, SSAStepper())

sol_array = zeros(4, length(sol))
sol_array[1:3, :] = Array(sol)
sol_array[4, :] = sum(sol_array[1:3, :], dims = 1);
```
```{julia}
f = Figure()
ax = Axis(
    f[1, 1],
    xlabel = "Time",
    ylabel = "Number"
)
